---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r srr-tags, eval=FALSE, echo=FALSE}
#' @srrstats {G1.2} Life cycle badge added in the README and statement on state
#'   of development added in the "Contributing" section.
```

# rcrisp

<!-- badges: start -->
[![R-CMD-check](https://github.com/CityRiverSpaces/rcrisp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CityRiverSpaces/rcrisp/actions/workflows/R-CMD-check.yaml)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.15793526.svg)](https://doi.org/10.5281/zenodo.15793526)
[![CRAN status](https://www.r-pkg.org/badges/version/rcrisp)](https://CRAN.R-project.org/package=rcrisp)
[![](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

rcrisp provides tools to automate the morphological delineation of riverside urban areas following the CRiSp (City River Spaces) method.

## Installation

You can install the released version of rcrisp from [CRAN](https://cran.r-project.org) with:

``` r
install.packages("rcrisp")
```

You can install the development version of rcrisp from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("CityRiverSpaces/rcrisp")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, warning=FALSE, message=FALSE}
library(rcrisp)

# Set location parameters
city_name <- "Bucharest"
river_name <- "Dâmbovița"

# Delineate river corridor
bd <- delineate(city_name, river_name, segments = TRUE)

# Plot delineation object
plot(bd)
```

## Contributing

rcrisp is in a stable state of development, with some degree of active subsequent development as envisioned by the primary authors.

We also look very much forward to contributions. See the [Contributing Guide](https://github.com/CityRiverSpaces/rcrisp/blob/main/.github/CONTRIBUTING.md) for further details.

This package is released with a [Contributor Code of Conduct](https://github.com/CityRiverSpaces/rcrisp/blob/main/.github/CODE_OF_CONDUCT.md). By contributing to this project you agree to abide by its terms.
