---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CRiSp

<!-- badges: start -->
[![R-CMD-check](https://github.com/CityRiverSpaces/CRiSp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CityRiverSpaces/CRiSp/actions/workflows/R-CMD-check.yaml)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

CRiSp (City River Spaces) provides tools to automate the morphological delineation of riverside urban areas.

## Installation

You can install the released version of CRiSp from [CRAN](https://cran.r-project.org) with:

``` r
# Install from CRAN
install.packages("CRiSp")
```

You can install the development version of CRiSp from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("CityRiverSpaces/CRiSp")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, warning=FALSE, message=FALSE}
library(CRiSp)

# Set location parameters
city_name <- "Bucharest"
river_name <- "Dâmbovița"
epsg_code <- 32635

# Get base layer for plotting
bb <- get_osm_bb(city_name)
streets <- get_osm_streets(bb, epsg_code)$geometry

# Delineate river corridor
corridor <- delineate(city_name, river_name)$corridor

# Plot results
plot(corridor, border = "orange", lwd = 3)
plot(streets, add = TRUE)
```

## Contributing

We look very much forward to contributions to the package. See the [Contributing Guide](.github/CONTRIBUTING.md) for further details.

This package is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project you agree to abide by its terms.
