---
title: "5. Corridor segmentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Corridor segmentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warnings = FALSE
)
```

```{r setup}
library(CRiSp)
```

For a more detailed analysis of an urban river corridor, corridor-level delineation may not be sufficient. The corridor needs to be subdivided into smaller morphological units. Segmentation is a process of subdividing the corridor by using major transversal road or rail infrastructure lines.

By default, the all-in-one function `delineate()` only returns the corridor boundary. The corridor can be segmented either by setting the argument `segments = TRUE` in `delineate()` or by using the `delineate_segments()` function in a separate step.

To demonstrate this as a separate step, we will use the `bucharest_delineation$corridor`, `bucharest_osm$streets` and `bucharest_osm$railways` layers from the package data as input.

We first prepare the network and select all the streets and railways that cover the river corridor plus a small buffer region (see also `vignette("network-preparation")`):

```{r network, warning=FALSE}
# Add a 100 meter buffer region around the corridor
corridor_buffer <- sf::st_buffer(bucharest_delineation$corridor, 500)

# Filter the streets and railwayas to the buffer area
streets <- bucharest_osm$streets |>
  sf::st_filter(corridor_buffer, .predicate = sf::st_intersects)
railways <- bucharest_osm$railways |>
  sf::st_filter(corridor_buffer, .predicate = sf::st_intersects)

# Build combined street and railway network
network_filtered_before <- rbind(streets, railways) |>
  as_network()

# Build combined street and railway network
network_filtered_after <- rbind(bucharest_osm$streets,
                                bucharest_osm$railways) |>
  as_network() |>
  tidygraph::activate("nodes") |>
  tidygraph::filter(sfnetworks::node_intersects(corridor_buffer)) |>
  # keep only the main connected component of the network
  tidygraph::activate("nodes") |>
  dplyr::filter(tidygraph::group_components() == 1)
```

We then delineate segments in the corridor. The algorithm spits the corridor using river-crossing transversal edges that form continuous lines in the network:

```{r segmentation, warning=FALSE}
segmented_corridor_1 <- delineate_segments(bucharest_delineation$corridor,
                                           network_filtered_before,
                                           bucharest_osm$river_centerline)
segmented_corridor_2 <- delineate_segments(bucharest_delineation$corridor,
                                           network_filtered_after,
                                           bucharest_osm$river_centerline)
plot(segmented_corridor_1, border = "orange", lwd = 3)
plot(segmented_corridor_2, add = TRUE, border = "blue")
```

