---
title: "4. Corridor delineation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{4. Corridor delineation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  eval = FALSE  #TODO remove this to enable code chunks after #87 is merged
)
```

```{r setup, warning=FALSE, message=FALSE}
library(CRiSp)
```

In this notebook we explore how to delineate an urban river corridor using river Dâmbovița in Bucharest, Romania's capital, as the study area. We will use OpenStreetMap (OSM) data, first from the Overpass API and then from a local file.

```{r variables}
city_name <- "Bucharest, Romania"   # Be specific and spell as in OSM
river_name <- "Dâmbovița"           # Spell as in OSM
```

We start by demonstrating the use of the all-in-one `delineate()` function which does the following three steps:

1. Fetches city boundary, street and rail network, as well as river centreline and surface data from the [Overpass API](https://wiki.openstreetmap.org/wiki/Overpass_API) as shown in `vignette("getting-osm-data")`;

2. Pre-processes the street and rail network for delineation as in `vignette("network-preparation")`

3. Constructs the initial corridor based on the chosen method. If the `method = "buffer"` is chosen, the corridor is constructed by buffering the river centreline and surface data with a given buffer distance. If the `method = "valley"` is chosen instead, the Cost Distance Accumulation algorithm is used to delineate the corridor based on the digital elevation model (DEM) of the area retrieved from the [Earth Search API](https://element84.com/earth-search/).

4. Delineates the corridor based on the pre-processed network and initial corridor. Optionally, the corridor is split into segments based on the network and the river space is delineated.

The `delineate()` function carries out the above steps in one and returns a list that by default contains the following output elements: `corridor`, `segments`, and `riverspace`.

```{r}
bucharest_dambovita <- delineate(
  city_name,
  river_name,
  segments = TRUE,
  riverspace = TRUE
)

# `$corridor` of the resulting object is an sf polygon representing the corridor
bucharest_dambovita$corridor

# `$segments` is an sf polygon representing the segments of the corridor
bucharest_dambovita$segments

# `$riverspace` is an sf polygon representing the delineated river space
bucharest_dambovita$riverspace

# All three elements of delineation
plot(bucharest_dambovita$corridor, border = NA)
plot(bucharest_dambovita$riverspace, col = "green", border = NA, add = TRUE)
plot(bucharest_osm$river_centerline, col = "blue", add = TRUE)
plot(bucharest_dambovita$segments, border = "lightblue", add = TRUE)
plot(bucharest_dambovita$corridor, border = "red", wt = 2, add = TRUE)
```

