---
title: "6. Study area around POI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{6. Study area around POI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(CRiSp)
library(sf)
```

## Introduction

In this article we explore how to delineate a study area around a point of interest (POI) using Bucharest as the study area. We choose a location along the River Dâmbovița. The POI is represented as a point in the WGS84 coordinate reference system.

```{r}
poi <- st_sfc(st_point(c(26.074123, 44.432875)), crs = 4326)
```

The `delineate_segment()` takes the POI as input. By default, it tries to identify the city and the river closest to it. Alternatively, it can take an already delineated corridor and street network as input. We will use the built-in data `bucharest_corridor` and `bucharest_streets`. It also takes care of any CRS transformations needed. Optionally, it can also generate the riverspace.

```{r}
segment_poi <- delineate_segment(poi, bucharest_corridor, bucharest_streets,
                                 riverspace = TRUE)
```

The output is a list object with three elements: `$poi`, `$segment` and `$riverspace`.

```{r}
plot(segment_poi$riverspace, col = "lightblue")
plot(segment_poi$segment, col = "blue", add = TRUE)
plot(segment_poi$poi, col = "red", size = 3, add = TRUE)
```

