---
title: "6. Corridor segmentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{6. Corridor segmentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warnings = FALSE
)
```

```{r setup}
library(rcrisp)
library(sf)

bucharest_osm <- get_osm_example_data()
bucharest_dem <- get_dem_example_data()

if (any(is.null(bucharest_osm), is.null(bucharest_dem))) {
  cat("NOTE: Example data was not found; ",
    "subsequent code chunks will be skipped.\n", sep = "")
  knitr::opts_chunk$set(eval = FALSE)
}
```

```{r srr-tags, eval=FALSE, echo=FALSE}
#' @srrstats {G2.10} This vignette uses `sf::st_geometry()` to extract the
#'   geometry column from the `sf` objects `bucharest_osm$river_centerline` and
#'   `streets`. This is used when only geometry information is needed from that
#'   point onwards and all other attributes (i.e., columns) can be safely
#'   discarded. The object returned by `sf::st_geometry()` is a simple feature
#'   geometry list column of class `sfc`.
#' @srrstats {SP2.2a} This vignette demonstrates the compatibility of `rcrisp`
#'   routines with `sf` workflows.
```

For a more detailed analysis of an urban river corridor, corridor-level delineation may not be sufficient. The corridor needs to be subdivided into smaller morphological units. Segmentation is a process of subdividing the corridor by using major transversal road or rail infrastructure lines.

By default, the all-in-one function `delineate()` only returns the corridor boundary. The corridor can be segmented either by setting the argument `segments = TRUE` in `delineate()` or by using the `delineate_segments()` function in a separate step.

To demonstrate this as a separate step, we will use the `bucharest_dambovita$corridor` from the package data, as well as `bucharest_osm$streets` and `bucharest_osm$railways` from rcrisp example data as input.

We first prepare the network and select all the streets and railways that cover the river corridor plus a small buffer region (see also `vignette("network-preparation")`):

```{r network, warning=FALSE}
# Add a buffer region around the corridor
corridor_buffer <- sf::st_buffer(bucharest_dambovita$corridor, 500)

# Filter the streets and railwayas to the buffer area
streets <- bucharest_osm$streets |>
  sf::st_filter(corridor_buffer, .predicate = sf::st_covered_by)
railways <- bucharest_osm$railways |>
  sf::st_filter(corridor_buffer, .predicate = sf::st_covered_by)

# Build combined street and railway network
network_filtered <- rbind(streets, railways) |>
  as_network()

```

We then delineate segments in the corridor. The algorithm spits the corridor using river-crossing transversal edges that form continuous lines in the network:

```{r segmentation, warning=FALSE}
segmented_corridor <-
  delineate_segments(bucharest_dambovita$corridor,
                     network_filtered,
                     st_geometry(bucharest_osm$river_centerline))
plot(st_geometry(streets))
plot(segmented_corridor, border = "orange", lwd = 3, add = TRUE)
```

