---
title: "8. Benchmarking across city-river cases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking across city-river cases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  message   = FALSE,
  warning   = FALSE,
  fig.width = 7,
  fig.height = 8,
  eval = requireNamespace("ggplot2", quietly = TRUE)
)
```

This vignette documents how `rcrisp` performs across a diverse set of
European city-river cases. Benchmarks were run on a single machine in a
single session using OpenStreetMap and Copernicus GLO-30 DEM data downloaded
live with the `get_osm()` and `get_dem()` functions, respectively. The timings
from these two data retrieval functions therefore include network latency and
reflect the volume of spatial data available for each case, not just CPU work.

```{r prepare-data-and-plot}
library(dplyr)
library(ggplot2)

# Load and prepare data
raw <- read.csv("../benchmark/output/timings.csv", stringsAsFactors = FALSE) |>
  mutate(case = paste0(city_name, " - ", river_name))

# Define pipeline steps
io_steps <- c("get_osm", "get_dem")
compute_steps <- c("delineate_valley", "as_network", "delineate_corridor",
                   "delineate_segments", "delineate_riverspace")
all_steps <- c(io_steps, compute_steps)

# Define display labels for steps
step_labels <- c(
  get_osm              = "get_osm()",
  get_dem              = "get_dem()",
  delineate_valley     = "delineate_valley()",
  as_network           = "as_network()",
  delineate_corridor   = "delineate_corridor()",
  delineate_segments   = "delineate_segments()",
  delineate_riverspace = "delineate_riverspace()"
)

# Define colours: blues for I/O steps, warm tones for computation
step_colours <- c(
  "get_osm()"              = "#2166AC",
  "get_dem()"              = "#74ADD1",
  "delineate_valley()"     = "#FDCC8A",
  "as_network()"           = "#FC8D59",
  "delineate_corridor()"   = "#E34A33",
  "delineate_segments()"   = "#B30000",
  "delineate_riverspace()" = "#7F0000"
)

# Prepare data for plotting
case_order <- raw |>
  filter(step == "total", status == "success") |>
  arrange(elapsed_s) |>
  pull(case)

bar_data <- raw |>
  filter(case %in% case_order, step %in% all_steps, status == "success") |>
  mutate(
    step_label = factor(step_labels[step], levels = step_labels[all_steps]),
    case = factor(case, levels = case_order)
  )

case_totals <- bar_data |>
  group_by(case) |>
  summarise(total = sum(elapsed_s), .groups = "drop")
```

## Run time by case and step

Blues indicate **data retrieval** functions (`get_osm()`, `get_dem()`); warm
colours indicate **computation** functions. Cases are sorted from fastest
(bottom) to slowest (top).

```{r bar-chart, fig.alt = "Horizontal stacked bar chart showing elapsed time per city-river case, broken down by pipeline step. Blues represent data retrieval functions and warm colours represent computation functions, stacked left to right in pipeline order."}
ggplot(bar_data, aes(x = elapsed_s, y = case, fill = step_label)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  geom_text(
    data = case_totals,
    aes(x = total, y = case, label = paste0(round(total), " s")),
    inherit.aes = FALSE,
    hjust = -0.15,
    size  = 3
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = step_colours,
                    guide  = guide_legend(nrow = 2)) +
  labs(x = "Elapsed time (s)", y = NULL, fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position    = "top",
    legend.key.size    = unit(0.5, "cm"),
    panel.grid.major.y = element_blank()
  )
```

## Key findings

- **`get_osm()` dominates total runtime**, typically accounting for 70–90% of
  wall-clock time. The wide range (from ~8 s for Plovdiv to ~355 s for Paris)
  reflects OSM data volume rather than algorithmic complexity.

- **`get_dem()` is fast and consistent** (1.8–7 s), since the number of DEM
  tiles depends on geographic extent, not urban density.

- **Among computation steps, `delineate_corridor()` and `delineate_segments()`
  show the most variability** relative to their mean times.
  `delineate_corridor()` is noticeably slower for large, dense street networks
  (Paris: 10.4 s vs. typical 1–2 s), while `delineate_riverspace()` is more
  consistently fast across all cases.

- **`as_network()` and `delineate_corridor()` scale with street network size**,
  and are noticeably slower for large, dense cities such as Paris
  (~26 s combined), Rome (~17 s), and Budapest (~7 s).

- **`delineate_valley()` and `delineate_segments()` are generally fast**
  across all cases, though a few complex cases exceed 5 s.
